Effects of the body force on the pedestrian and evacuation dynamics

### 2020-02-05 paper.tex
cambie los <> de los grficos

Fig.~\ref{venn_3parameters} provides a picture of the parameters' relations 
after introducing $\mathcal{K}_c$. Surprisingly, $\mathcal{K}_c$ appears as a
wider set (say, a four elements set) than $\mathcal{A}$ or $\mathcal{K}$ (three 
elements' sets). It shares the parameter $v_d$ with $\mathcal{A}$ and the 
parameter $B$ with $\mathcal{K}$. The practical consequence to these (logical)
relations is that $v_d$ or $B$ affect simultaneously two ``control'' parameters 
of the collective dynamics. Conversely, either $v_d$ and $B$ may counterbalance 
$\mathcal{K}_c$ in order to keep the colletive dynamic (qualitatively) 
unchanged.  \\

@@

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_bottleneck.eps}}\\
\caption[width=0.47\columnwidth]{Mean velocity in the $x-$direction as a function 
of the stiffness level $k_n$ for different desired velocities (see label). The 
data corresponds to a bottleneck with periodic boundary conditions (re-injecting 
pedestrians). The average was taken every five seconds once the crowd reaches the 
stationary state ($t=$20~s) until the end of the simulation ($t=$110~s). }
\label{kn_vs_vx_bottleneck}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the pedestrians desired velocity. Each symbol indicates the $k_n$ value corresponding to the body force (see the labels). The data corresponds to a bottleneck with periodic boundary conditions (re-injecting pedestrians). The average was taken every five seconds once the crowd reached the stationary state (say, $t=$20~s) until the end of the simulation ($t=$110~s).}
\label{degree_overlap_vd}
\end{figure}

The regime for $v_d<2\,$m/s is qualitatively different from the clogged regime since most of the 
pedestrians are not touching each other. The only pedestrians who touch each 
other are the ones who are being re-injected on the opposite side of the room.
This collides against the bulk, providing some kind of overlap during very short time. We will
not analize this regime.\\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Snapshots of the contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)$~m, the width of the door is 0.92~m (equivalent to 2 pedestrian\textsc{\char13}s diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). (a) and (b) correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. (c) and (d) correspond to simulations with $k_n$=1.2\,E6 with $v_d=$2 and $v_d=$10 respectively.}
\label{network_bottleneck}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./pbc_vs_vd_multi_kn.eps}
\caption{\label{pbc_vs_vd_multi_kn} Blocking cluster probability as a function of $v_d$ for different stiffness levels (see label). The probability is calculated as the amount of time a blocking cluster is present divided by the overall simulation time. The situation corresponds to a bottleneck with 225 pedestrians under periodic boundary conditions (re-injection of pedestrians once they left the room). The sampled interval was set to $t_f =$ 1000~s. }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_dens_multi_kn.eps}\label{degree_dens}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_dens_multi_kn.eps}\label{overlap_dens}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the global density for different $k_n$ values. (b) mean overlap as a function of the global density. The mean values are averages over all the pedestrians and over time once the system reaches the stationary state. Both measurements correspond to corridor geometry with desired velocity $v_d=1$.}
\label{network_corridor}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor. The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2$~E5.}
\label{network_corridor}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./triangles.eps}
\caption{\label{triangles} Triangles per node as a function of the global density. A triangle in a network is defined as three nodes all connected. Each marker corresponds to a different stiffness $k_n$ (see the label in the plot). The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$.  }
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_corridor.eps}
\caption{\label{kn_vs_vx_corridor} Mean velocity in the longitudinal coordinate $(v_x)$ as a function of the stiffness $k_n$ for three different global densities (see label in the plot). The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$.      }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./vx_profile.eps}\label{vx_profile}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./strain_rate_vs_kn.eps}\label{strain_rate_vs_kn}}\\
\caption[width=0.47\columnwidth]{(a) Velocity in the $x$ coordinate as a function of the transversal coordinate $y$ for three different stiffness values $k$ (see the label). (b) Strain rate as a function of the stiffness level $k_n$ for three different global density values (see label). The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. The global density was 7 p~m$^{-2}$.  }
\label{profile_strain}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./flow-density_multifric_nobodyforce.eps}\label{flow_density_no_body_force}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./flow-density_multifric_bodyforce.eps}\label{flow_density_body_force}}\\
\caption[width=0.47\columnwidth]{Flow vs. global density. The flow is calculated in a circular area of R=1~m at the center of the corridor. The circular markers correspond to the original friction of the SFM, the "+" symbol corresponds to the friction increased by a factor of five and the triangles correspond to the friction increased by a factor of ten. The desired velocity was $v_d=1$ in all the cases. (a) corresponds to simulations without body force ($\mathcal{K}_c =$0) and (b) corresponds to a body force with the original value of the body stiffness ($\mathcal{K}_c =$68).}
\label{flow_density}
\end{figure}

>>>>

Fig.~\ref{venn_3parameters} provides a picture of the parameters' relations 
after introducing $\mathcal{K}_c$. Surprisingly, $\mathcal{K}_c$ appears as a
wider set (say, a four elements set) than $\mathcal{A}$ or $\mathcal{K}$ (three 
elements' sets). It shares the parameter $v_d$ with $\mathcal{A}$ and the 
parameter $B$ with $\mathcal{K}$. The practical consequence to these (logical)
relations is that $v_d$ or $B$ affect simultaneously two ``control'' parameters 
of the collective dynamics. Conversely, either $v_d$ and $B$ may counterbalance 
$\mathcal{K}_c$ in order to keep the collective dynamic (qualitatively) 
unchanged.  \\

@@

\subsection{\label{blocking_clusters} Blocking clusters}

A characteristic feature of pedestrian dynamics is the formation of clusters. Clusters of pedestrians can 
be defined as the set of individuals that for any member of the group (say, $i$) there exists at least another member belonging to the same group ($j$) in contact with the former. Thus, we define a ``granular cluster'' ($C_g$) following the mathematical formula given in Ref.~\cite{dorso_2005}

\begin{equation}
C_g:P_i~\epsilon~ C_g \Leftrightarrow \exists~ j~\epsilon~C_g / r_{ij} < (R_i+R_j) \label{ec-cluster}
\end{equation}

where ($P_i$) indicate the \textit{ith} pedestrian and $R_i$ is his (her) radius (shoulder width). That means, $C_g$ is a
set of pedestrians that interact not only with the social force, but also with physical forces (\textit{i.e.} friction force and body force).
A ``blocking cluster'' is defined as the subset of clusterized particles (granular cluster) closest to the door whose first 
and last component particles are in contact with the walls at both sides of the door. This clogging structure is responsible for 
worsening the evacuation performance.

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_bottleneck.eps}}\\
\caption[width=0.47\columnwidth]{Mean velocity in the $x-$direction as a function 
of the stiffness level $k_n$ for different desired velocities (see label). The 
data corresponds to a bottleneck with periodic boundary conditions (re-injecting 
pedestrians). The average was taken every five seconds once the crowd reaches the 
stationary state ($t=$20~s) until the end of the simulation ($t=$110~s). 
Color online only. }
\label{kn_vs_vx_bottleneck}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_vd_multi_kn.eps}\label{degree_vd}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_vd_multi_kn.eps}\label{overlap_vd}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the pedestrian’s desired velocity (m/s). (b) Mean overlap as a function of the pedestrians desired velocity. Each symbol indicates the $k_n$ value corresponding to the body force (see the labels). The data corresponds to a bottleneck with periodic boundary conditions (re-injecting pedestrians). The average was taken every five seconds once the crowd reached the stationary state (say, $t=$20~s) until the end of the simulation ($t=$110~s). Color online only.}
\label{degree_overlap_vd}
\end{figure}

The regime for $v_d<2\,$m/s is qualitatively different from the clogged regime since most of the 
pedestrians are not touching each other (as was already observed in Ref~\cite{dorso_2005,dorso_2011}). The only pedestrians who touch each 
other are the ones who are being re-injected on the opposite side of the room.
This collides against the bulk, providing some kind of overlap during very short time. We will
not analize this regime.\\

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn0.eps}\label{network_vd2_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn0.eps}\label{network_vd10_kn0}}\\
        \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd2_kn1200000.eps}\label{network_vd2_kn1200000}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_vd10_kn1200000.eps}\label{network_vd10_kn1200000}}\\
\caption[width=0.47\columnwidth]{Snapshots of the contact networks of a 225 pedestrian evacuation through a bottleneck. The door is placed at $(x,y)=(20,10)$~m, the width of the door is 0.92~m (equivalent to 2 pedestrian\textsc{\char13}s diameter). The lines that connect the nodes (pedestrians) represent the contact between them. The color represents the degree (the number of pedestrians with which it is connected). (a) and (b) correspond to a simulation without body force with $v_d=$2 and $v_d=$10 respectively. (c) and (d) correspond to simulations with $k_n$=1.2\,E6 with $v_d=$2 and $v_d=$10 respectively. Color online only.}
\label{network_bottleneck}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./pbc_vs_vd_multi_kn.eps}
\caption{\label{pbc_vs_vd_multi_kn} Blocking cluster probability as a function of $v_d$ for different stiffness levels (see label). The probability is calculated as the amount of time a blocking cluster is present divided by the overall simulation time. The situation corresponds to a bottleneck with 225 pedestrians under periodic boundary conditions (re-injection of pedestrians once they left the room). The sampled interval was set to $t_f =$ 1000~s. Color online only. }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./degree_vs_dens_multi_kn.eps}\label{degree_dens}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./overlap_vs_dens_multi_kn.eps}\label{overlap_dens}}\\
\caption[width=0.47\columnwidth]{(a) Mean degree as a function of the global density for different $k_n$ values. (b) mean overlap as a function of the global density. The mean values are averages over all the pedestrians and over time once the system reaches the stationary state. Both measurements correspond to corridor geometry with desired velocity $v_d=1$. Color online only.}
\label{network_corridor}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_kn0.eps}\label{network_d6_kn0}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./network_d6_knE5.eps}\label{network_d6_knE5}}\\
\caption[width=0.47\columnwidth]{Contact network of the pedestrians along the corridor. The lines that connect the nodes (pedestrians) represent the contact between them. The colors stand for the degree of the node (the number of pedestrians that are in contact with him/her). The corridor was 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. (a) corresponds to a simulation without body force and (b) corresponds to a simulation with $k_n=1.2$~E5. Color online only.}
\label{network_corridor}
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./triangles.eps}
\caption{\label{triangles} Triangles per node as a function of the global density. A triangle in a network is defined as three nodes all connected. Each marker corresponds to a different stiffness $k_n$ (see the label in the plot). The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. Color online only. }
\end{figure}

@@

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]{./kn_vs_vx_corridor.eps}
\caption{\label{kn_vs_vx_corridor} Mean velocity in the longitudinal coordinate $(v_x)$ as a function of the stiffness $k_n$ for three different global densities (see label in the plot). The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. Color online only.      }
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./vx_profile.eps}\label{vx_profile}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./strain_rate_vs_kn.eps}\label{strain_rate_vs_kn}}\\
\caption[width=0.47\columnwidth]{(a) Velocity in the $x$ coordinate as a function of the transversal coordinate $y$ for three different stiffness values $k$ (see the label). (b) Strain rate as a function of the stiffness level $k_n$ for three different global density values (see label). The measurements correspond to a corridor of 28~m $\times$ 22~m with periodic boundary conditions and $v_d=1$. The global density was 7 p~m$^{-2}$. Color online only. }
\label{profile_strain}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./flow-density_multifric_nobodyforce.eps}\label{flow_density_no_body_force}}\ 
    \subfloat[]{\includegraphics[width=0.49\columnwidth]{./flow-density_multifric_bodyforce.eps}\label{flow_density_body_force}}\\
\caption[width=0.47\columnwidth]{Flow vs. global density. The flow is calculated in a circular area of R=1~m at the center of the corridor. The circular markers correspond to the original friction of the SFM, the "+" symbol corresponds to the friction increased by a factor of five and the triangles correspond to the friction increased by a factor of ten. The desired velocity was $v_d=1$ in all the cases. (a) corresponds to simulations without body force ($\mathcal{K}_c =$0) and (b) corresponds to a body force with the original value of the body stiffness ($\mathcal{K}_c =$68). Color online only.}
\label{flow_density}
\end{figure}
